Hydra battles in Coq

### 2020-09-03 latex/chapter-schutte.tex
table of notations
minor corrections

\index{Types!Ord@Ord (countable ordinals)}

@@

@@

Let us import the library \texttt{Coq.Logic.Epsilon}, which contains the following axiom.

@@

\index{Functions!iota description operator}

@@

\index{Functions!succ (in type Ord)}

@@

@@

\index{Notations!omega@omega (the least infinite ordinal)}

\begin{Coqsrc}
Definition _omega := omega_limit finite.

@@

\index{Functions!phi0@phi0 (exponential of base omega)}

>>>>

\label{types:Ord}

@@

\index{Coq!Type classes}
\label{types:WO}

@@

Let us import the library \texttt{Coq.Logic.Epsilon}, which contains the following axiom and lemmas.

@@

@@

\label{Functions:succ-sch}

@@

\label{sect:notation-F-sch}

@@

\label{sect:notation-omega}
\begin{Coqsrc}
Definition _omega := omega_limit finite.

@@

### 2020-09-03 latex/chapter-termination.tex
minor corrections

\index{Types!T1@T1 (ordinal terms below $\epsilon_0$)}
\index{Constants!zero:T1}

@@

\index{Types!T2@T2 (ordinals terms in Veblen normal form)}
\emph{Please look also at the library
\href{../src/html/hydras.Gamma0.T2.html}{Gamma0.T2.html}}.

@@

For representing finite ordinals, \emph{i.e.} natural numbers, we first introduce a notation for terms of the form $n+1$, then define a coercion from type \texttt{nat} into \texttt{T1}.

@@

@@

@@

\index{Notations!omega@omega (the least infinite ordinal)}
\begin{Coqsrc}
Notation omega := (ocons (ocons zero 0 zero) 0 zero): t1_scope.
\end{Coqsrc}

@@

\subsection{Making normality implicit}
  We would like to get rid of terms of type \texttt{T1} which are not in Cantor normal form.
A simple way to do this is to consider statements of the form 
\texttt{forall alpha: T1, nf alpha -> $P$ alpha}, where $P$ is a predicate over type \texttt{T1}, lile in the following lemma.

@@

\label{sect:E0-def}
\index{Types!E0@ EO (ordinal terms below $\epsilon_0$ in normal form)}

@@

Equality in \texttt{E0} is just Leibniz equality. Note that, since \texttt{nf} is
defined by a Boolean function, for  any term $\alpha:\texttt{T1}$, there exists at most one proof of \texttt{nf $\alpha$}, thus two ordinals of type \texttt{E0} are
equal if and only iff their projection to \texttt{T1} are equal.

@@

\index{Notations!omega@omega (the least infinite ordinal)}
\index{Constants!zero:T1}

@@

\index{Functions!T1.succ}

@@

@@

\index{Types!ppT1@ppT1 (pretty printed ordinal terms)}
\index{Functions!pp@ pp (pretty printing terms in Cantor normal form)}

>>>>

\label{types:T1}
\index{Constants!zero:T1}

@@

\label{types:T2}
\emph{Please look also at the library
\href{../src/html/hydras.Gamma0.T2.html}{Gamma0.T2.html}}.

@@

For representing finite ordinals, \emph{i.e.} natural numbers, we first introduce a notation for terms of the form $n+1$, then define a coercion from type \texttt{nat} into \texttt{T1}.
\label{sect:notation-FS}

@@

\label{sect:notation-F}

@@

Notation "'F' n" := (fin n) (at level 10) : t1_scope.

@@

\label{sect:omega-notation2}
\begin{Coqsrc}
Notation omega := (ocons (ocons zero 0 zero) 0 zero): t1_scope.
\end{Coqsrc}

@@

\subsection{Making normality implicit}
  We would like to get rid of terms of type \texttt{T1} which are not in Cantor normal form.
A simple way to do this is to consider statements of the form 
\texttt{forall alpha: T1, nf alpha -> $P$ alpha}, where $P$ is a predicate over type \texttt{T1}, like in the following lemma \footnote{Ordinal addition is formally defined a little later (page~\ref{sect:infix-plus-T1})}.

@@

\label{sect:E0-def}
\label{types:E0}

@@

Equality in \texttt{E0} is just Leibniz equality. Note that, since \texttt{nf} is
defined by a Boolean function, for  any term $\alpha:\texttt{T1}$, there exists at most one proof of \texttt{nf $\alpha$}, thus two ordinals of type \texttt{E0} are
equal if and only iff their projection to \texttt{T1} are equal (see also Sect.~\vref{sect:eq-proof-unicity}).

@@

\label{sect:omega-T1}
\index{Constants!zero:T1}

@@

\label{Functions:succ-T1}

@@

\label{sect:infix-plus-T1}

@@

\label{types:ppT1}
\index{Functions!pp@ pp (pretty printing terms in Cantor normal form)}

### 2020-09-03 latex/hydras.tex
minor corrections

\begin{table}[h]
  \centering
  \begin{threeparttable}
    \caption{Main types and type classes}
\begin{tabular}{|c|c|c|l|}
\hline
Name & Description& Page \\\hline
Battle & Type class of battle types & \pageref{types:Battle} \\
E0 & (Well formed) ordinals below $\epsilon_0$ & \pageref{types:E0}\\
Hvariant & variants for proving terminatio of hydra battles & \pageref{sect:hvariant-def} \\
Hydra & tree-like representation of hydras & \pageref{types:Hydra} \\
Hydrae & finite sequences of hydras & \pageref{types:Hydrae} \\
ON & ordinal  notations & \pageref{types:ON} \\
ON\_for & comparison of an ordinal notation with Schütte's model &  \pageref{types:ON-for}\\
ON\_iso & isomorphism of ordinal notations  & \pageref{types:ON-iso}\\

@@

\begin{table}[h]
  \centering
  \begin{threeparttable}
    \caption{Main functions}
\begin{tabular}{|r | c|c|c|c|l|}
\hline
Name & Gallina & Math & Description & Page \\ \hline
canon & canon alpha i & $\canonseq{\alpha}{i}$& Canonical sequence & \pageref{Functions:canon}\\
canonS & canonS alpha i & $\canonseq{\alpha}{i+1}$& Helper for canon & \pageref{Functions:canonS}\\
F\_ & F\_ alpha n & $F_\alpha(n)$ & Wainer's \texttt{F} fast growing hierarchy & \pageref{Functions:F-alpha} \\
iterate & iterate $f$ $n$ $x$ & $f^{(n)}(x)$ & Functional iteration & \pageref{Functions:iterate} \\
\hline
\end{tabular}
\end{threeparttable}
\end{table}

\vspace{4pt}
\begin{table}[h]
  \centering
  \begin{threeparttable}
    \caption{Infix notations}
\begin{tabular}{|c|c|c|c|l|}
\hline
Name & Gallina&Math& Description& Page \\\hline
 \texttt{on\_lt} & \texttt{alpha o< beta}&$\alpha<\beta$& ordinal inequality \tnote{1} & \pageref{sect:on-lt-notation}\\
 \texttt{on\_le} & \texttt{alpha o<= beta}&$\alpha\leq\beta$& ordinal inequality & \pageref{sect:on-lt-notation}\\
\texttt{plus} & \texttt{alpha + beta} & $\alpha + \beta$ & Ordinal addition & \pageref{sect:infix-oplus} \\
\texttt{oplus} & \texttt{alpha o+ beta} & $\alpha \oplus \beta$ & Hessenberg sum & \pageref{sect:infix-oplus} \\
\texttt{round} & \texttt{h -1-> h'} & & one round of a battle & \pageref{sect:infix-round} \\
\texttt{rounds} & \texttt{h -+-> h'} & & one or more  rounds of a battle & \pageref{sect:infix-rounds} \\
\texttt{round\_star} & \texttt{h -*-> h'} & & any number of rounds of a battle & \pageref{sect:infix-rounds} \\

>>>>

\begin{table}[h]
  \centering
  \begin{threeparttable}
    \caption{Main types and type classes}
\begin{tabular}{|c|c|c|l|}
\hline
Name & Description& Page \\\hline
Battle & Type class of battle types & \pageref{types:Battle} \\
E0 & (Well formed) ordinals below $\epsilon_0$ & \pageref{types:E0}\\
Hvariant & variants for proving termination of hydra battles & \pageref{sect:hvariant-def} \\
Hydra & tree-like representation of hydras & \pageref{types:Hydra} \\
Hydrae & finite sequences of hydras & \pageref{types:Hydrae} \\
ON & ordinal  notations & \pageref{types:ON} \\
ON\_for & comparison of an ordinal notation with Schütte's model &  \pageref{types:ON-for}\\
ON\_iso & isomorphism of ordinal notations  & \pageref{types:ON-iso}\\

@@

\begin{table}[h]
  \centering
  \begin{threeparttable}
    \caption{Main functions}
\begin{tabular}{|r | c|c|c|c|l|}
\hline
Name & Gallina & Math & Description & Page \\ \hline
canon & canon alpha i & $\canonseq{\alpha}{i}$& Canonical sequence & \pageref{Functions:canon}\\
canonS & canonS alpha i & $\canonseq{\alpha}{i+1}$& Helper for canon & \pageref{Functions:canonS}\\
F\_ & F\_ alpha n & $F_\alpha(n)$ & Wainer's \texttt{F} fast growing hierarchy & \pageref{Functions:F-alpha} \\
H\_ & H\_ alpha n & $H_\alpha(n)$ & Hardy's \texttt{H} fast growing hierarchy & \pageref{Functions:H-alpha} \\
iterate & iterate $f$ $n$ $x$ & $f^{(n)}(x)$ & Functional iteration & \pageref{Functions:iterate} \\
L\_ & L\_ alpha k & $L_\alpha(k)$ & final step of a standard path & \pageref{Functions:L-alpha} \\
succ & succ alpha & & Successor & \pageref{Functions:succ-T1}, \pageref{Functions:succ-sch} \dots \\
\hline
\end{tabular}
\end{threeparttable}
\end{table}

\vspace{4pt}
\begin{table}[h]
  \centering
  \begin{threeparttable}
    \caption{Infix notations}
\begin{tabular}{|c|c|c|c|l|}
\hline
Name & Gallina&Math& Description& Page \\\hline
 \texttt{on\_lt} & \texttt{alpha o< beta}&$\alpha<\beta$& ordinal inequality \tnote{1} & \pageref{sect:on-lt-notation}\\
 \texttt{on\_le} & \texttt{alpha o<= beta}&$\alpha\leq\beta$& ordinal inequality & \pageref{sect:on-lt-notation}\\
\texttt{plus} & \texttt{alpha + beta} & $\alpha + \beta$ & ordinal addition & \pageref{sect:infix-plus-T1} ,\dots\\
\texttt{oplus} & \texttt{alpha o+ beta} & $\alpha \oplus \beta$ & Hessenberg sum & \pageref{sect:infix-oplus} \\
\texttt{round} & \texttt{h -1-> h'} & & one round of a battle & \pageref{sect:infix-round} \\
\texttt{rounds} & \texttt{h -+-> h'} & & one or more  rounds of a battle & \pageref{sect:infix-rounds} \\
\texttt{round\_star} & \texttt{h -*-> h'} & & any number of rounds of a battle & \pageref{sect:infix-rounds} \\
